---
title: "US Birth Data"
author: "Sindre Sønvisen (sso149)"
date: "10/29/2020"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, comment=NA, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("read_files.r")
```

This pages shows some statistics regarding child birth in US year 2017, 2018 and 2019. The data is provided by: Centers for Disease Control and Prevention (CDC). Link to the data used: [linked phrase](https://www.cdc.gov/nchs/data_access/vitalstatsonline.htm).

```{r data, echo=FALSE, comment=NA, message=FALSE }
# ALl the data and graph code is in read_files.r
```

# Births by Gender
Below is a bar graph of the number of girls proportional to boys born each year.
```{r plot gender, echo=FALSE, comment=NA, message=FALSE }
graph_sex #from read_files.r
```

# Weight by gender:
Below is a graph of the average weight for both genders each year.
```{r plot weight_by_gender, echo=FALSE, comment=NA, message=FALSE }
graph_weight_by_sex #from read_files.r
```

# Gender by weekday:
Below is a bar graph of the proportional of each gender born on the different days of the week each year.
```{r plot gender_by_weekday, echo=FALSE, comment=NA, message=FALSE }
graph_sex_by_weekday #from read_files.r
```

### Method used to pars and present the data:
From tidyverse map is used with read_fwf to read and extract the data needed from all the data files, afterwards bind_rows is used to combine all all the data in to one data frame. ggplot is used to make the bar graphs.



