---
title: "US Birth Data"
author: "Sindre Sønvisen (sso149)"
date: "10/29/2020"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, comment=NA, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("read_files.r")
```

This pages shows some statistics regarding child birth in US year 2017, 2018 and 2019. The data is provided by: Centers for Disease Control and Prevention (CDC). Link to the data used: [linked phrase](https://www.cdc.gov/nchs/data_access/vitalstatsonline.htm).

```{r data, echo=FALSE, comment=NA, message=FALSE }
#The data need to be unzipped and placed in a folder named data, or change the loc and or pattern argument bellow.
df <- read_files(loc="./data", pattern=".txt")
```

# Births by Gender
Below is a bar graph of the number of girls proportional to boys born.
```{r plot gender, echo=FALSE, comment=NA, message=FALSE }
bar_sex <- ggplot(df, aes(x = sex)) +
  geom_bar(stat = "count",
           width = 0.7,
           fill = c("red", "blue")) +
  ggtitle("Number of girls and boys born") +
  ylab("Number born") + xlab("Gender")
bar_sex
```

# Weight by gender:
Below is a bar graph of the average weight for both genders.
```{r plot weight_by_gender, echo=FALSE, comment=NA, message=FALSE }
bar_weight_by_sex <-
  ggplot(df[df$weight != 9999,], aes(x = sex, y = weight)) +
  geom_bar(
    stat = "summary",
    fun = "mean",
    width = 0.7,
    fill = c("red", "blue")
  ) +
  ggtitle("weight by gender") +
  ylab("Average weight in grams") + xlab("Gender")
bar_weight_by_sex
```

# Gender by weekday:
Below is a bar graph of the number og each gender born on the different days of the week.
```{r plot gender_by_weekday, echo=FALSE, comment=NA, message=FALSE }
#Make a new field with the day names.
days <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
df <- mutate(df, weekday=factor(days[week_day]))

bar_sex_by_weekday <-
  ggplot(df, aes(x = reorder(weekday, week_day), fill = sex)) +
  geom_bar(stat = "count",
           width = 0.7,
           position = position_dodge()
           ) +
  ggtitle("Gender by Weekday") +
  ylab("Number born") + xlab("Weekday") +
  labs(fill = "Gender")
bar_sex_by_weekday
```

### Method used to pars and present the data:
From tidyverse map is used with read_fwf to read and extract the data needed from all the data files, afterwards bind_rows is used to combine all all the data in to one data frame. ggplot is used to make the bar graphs.



